\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}
\begin{document}
Higher Certificate, Paper I, 2001. Question 8
yi =α + β xi +ε i
where y is the response (observation) and x the value of the explanatory variable on
that unit; { } i
ε is a set of independent, identically distributed random variables with
mean 0 and the same varianceσ 2 . Usually they are assumed Normal as a basis for
inference. i x is assumed "fixed", not "random".
(i) (a) There is an increasing trend, and the relationship between y and x
appears curvilinear.
(b) In simple regression R2 is the square of the correlation, r, between x
and y. In general it is the proportion of the variance of y which can be
explained by the dependence of y on all explanatory variables { } i x in the
model; hence it is the square of the correlation between ˆy and y.
In the ANOVA , regression SS 269.33 0.803, or 80.3%
total SS 335.37
= = .
(c) As x (% operating capacity) increases by 1 unit so y (profit) increases
by 0.31562 units.
A 95% confidence interval is 10 0.31562 ± t ×0.04942 , which is
0.31562 ± 0.11011or (0.2055, 0.4257) .
(d) Values of profit have been predicted for capacity 25%, 50% and 75%.
The confidence intervals for these predictions are those given; but note that
25% is far outside the range of available data (hence the remark about extreme
x values). A 95% confidence interval is an interval which should cover the
true y at a given x with probability 0.95, based on the fitted linear regression.
(ii) (a) The logarithmic plot shows that a linear regression in these units is a
much better fit. There is still an increasing trend.
(b) log10(profits) = −0.519 + 0.0177(capacity)
i.e. profits =10−0.519+0.0177(capacity) .
(c) For capacity = 25, the 95% limits are –0.2731 and +0.1210, and the
actual prediction is –0.0760, in log10 units. Anti-logging these (i.e. raising 10
to these powers) we find the 95% limits are 0.5332 and 1.3213.
The 'prediction' is 0.8395.
These limits do not overlap the limits on the previous model.
The prediction now is for a small profit, compared with a loss on the previous
model.
(iii) The scatter plots indicate that the logarithmic model is preferred, and so do the
plots of residuals which show a random pattern (as compared with a
systematic, curved one for the previous model). R2 also higher (92.4%) on the
log model.
But a log model cannot predict negative profits – i.e. losses – which are quite
possible in general though not for these data if used within the range of x
values given.
Extrapolation down to 25% is well outside the data and so is not reliable on
any model.
\end{document}
