\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}
\begin{document}
Higher Certificate, Paper I, 2001. Question 2
\begin{table}[ht!]
 \centering
 \begin{tabular}{|p{15cm}|}
 \hline
\noindent Text
\\ \hline
  \end{tabular}
\end{table}



%==========================================================================================%
2. (a) Three married couples at a dinner party sit down at random at a circular
table set for six people. Find the probability that
(i) each man sits between two women,

(ii) all three men sit together,

(iii) exactly two men sit together.

(b) A disease, which can only be diagnosed with certainty after death, exists in
a proportion p0 of the population. A clinical test is known such that
P(test is positive given disease is present) = p1
and
P(test is negative given disease is absent) = p2 .
(i) Find, in terms of p0, p1 and p2, the probability that a randomly
chosen individual who tests positive actually has the disease.

(ii) Calculate the answer to (i) in the case p0 = 0.005, p1 = 0.95,
p2 = 0.95, and comment on your result.
\end{framed}
\begin{enumerate}
\item (a) Fix the position of M1 (suppose him to be the host).
M1 Label the positions clockwise.
(1)
(6) (2) M2, M3, W1, W2, W3 may be arranged in 5! =
120 ways.
(5) (3)
(4)
   \item M2, M3 must occupy (3) and (5); W1, W2, W3 may occupy the other
places in 3! ways, making $2×3!$ arrangements. The probability is then
\[ \frac{12}{120} =\frac{1}{10}\]
   \item M2, M3 must occupy (2) and (6) or (2) and (3) or (5) and (6). In each
case, M2, M3 can be placed in two orders, making 6 positions altogether for
the three men. The women may again fill the remaining places in 3! ways.
The probability is 
\[ \frac{6 \times 6}{120} =\frac{36}{120} = \frac{3}{10}\]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
 \centering
 \begin{tabular}{|p{15cm}|}
 \hline
\noindent Text
\\ \hline
  \end{tabular}
\end{table}
   \item EITHER 1 1 3 3
10 10 5
− − = , because this is the only other arrangement
possible besides (i) and (ii);
OR by having M2 in (2), M3 in (4) or (5); M2 in (6), M3 in (3) or (4); M2 in
(3), M3 in (4); M2 in (4), M3 in (5); or any of these with M2, M3 interchanged,
giving 12 positionings of the men. There are again 3! orders for the women,
so the probability is 
\[ \frac{6 \times 12}{120} =\frac{72}{120} = \frac{3}{5}\]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
 \centering
 \begin{tabular}{|p{15cm}|}
 \hline
\noindent Text
\\ \hline
  \end{tabular}
\end{table}
\item Event D is "has disease", T is "tests positive".

\begin{framed}

\[ {P(D|T) = \frac{P(D \cap T)}{P(T)}\]

Remark 
\[ P(D \cap T)  = P(D|T) \times P(D) = P(T|D) \times P(T)\]

\[ P(D|T) = \frac{P(D \cap T)}{P(T)} = \frac{P(D|T) \times P(D)}{P(T)}\]

\end{framed}

Total Probability
\[P(T) = P(T \cap D) + P(T \cap D^{C})\]


\begin{eqnarray*}
P(D|T) &=&   \frac{P(D|T) \times P(D)}{P(T)}\\
&=& \frac{P(D|T) \times P(D)}{P(T \cap D) + P(T \cap D^{C})}\\
&=& \frac{P(D|T) \times P(D)}{P(T|D) \times P(D) + P(T|D) \times P(D^{C})}\\
=& \frac{p_1 \times p_0}{\left[p_1 \times p_0\right] + \left[(1- p_2 )(1- p_0)\right]}\\
\end{eqnarray*}



\begin{table}[ht!]
 \centering
 \begin{tabular}{|p{15cm}|}
 \hline
\noindent Text
\\ \hline
  \end{tabular}
\end{table}

\item 

\[\frac{0.95 0.005 }{(0.95 0.005) (0.05 0.995) } = \frac{0.00475 }{0.0545} = 0.0872.\]

The error rates in the clinical tests are large compared to the chance of having
the disease, so the calculated probability is very small.
\end{enumerate}

\end{document}
