\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


Higher Certificate, Paper III, 2001. Question 7

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[ht!]
     


\centering
     


\begin{tabular}{|p{15cm}|}
     


\hline 

7. (a) (i) Explain what is meant by the non-response problem in a sample survey, giving examples of where it might arise.  
 
  (ii) What steps might be taken to reduce the level of non-response?  
 
 
(b) A survey organisation was interested in seeing what proportions of managing directors of companies in different sectors
 of business expected prospects for their company to improve in the next six months.  
The results of a survey of two sectors are given in the following table. 
 
  Improve    No Yes Total Light engineering 58 67 125 Sector Banking and financial services 74 126 200 
 
(i) Calculate an approximate 99% confidence interval for the difference in proportions.  
 
(ii) Do the results suggest that the two sectors have the same views?  Give reasons for your answer.  
 
 
 
\\ \hline



\end{tabular}
    


\end{table}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{enumerate}[(a)]
\item  Non-response is a problem introduced by people refusing to reply to a
survey (or being genuinely unavailable), because they may be different in
some ways from those who do reply. People may not have the information to
reply to questions, may have different working or leisure habits, may be away
more frequently, may live in shared accommodation which is less easy to
locate, may be of one particular age-group.
For example, those who play sports would be more likely to be out evenings
or weekends, but their views on facilities would be different from those who
do not.
(ii) (1) Pilot testing of questionnaires to check clarity,
understandability, avoid giving offence by wording or by
including sensitive questions.
(2) Use interviewers who are well-trained, understand the aim and
purpose of the survey, and the meaning of questions, and are
able to put people at their ease.
(3) Give advance notice where appropriate, e.g. to the area or
group of people being surveyed.
(4) Revisit those unable to be interviewed through absence.
(b) Light engineering: n = 125, P(improve) = 67 0.536
125
= .
Banking & finance: n = 200, P(improve) = 126 0.630
200
= .
(i) Using a Normal approximation (1 )
,
p p
p
n
 − 
 
 
for each proportion, the
variance of their difference is ( ) ( ) 1 1 2 2 3
1 2
1 1
3.1551 10
p p p p
n n
− − −
+ = × .
An approximate 95% confidence interval for the true difference ( ) 2 1 π −π is
( ) 3
2 1 p − p \pm 1.96 3.1551×10− , i.e. (−0.016, 0.204) .
(ii) Strictly we cannot say they are different because this interval
includes 0; but the lower limit is only just below 0 so we might investigate
further, if possible, using larger samples.
\end{enumerate}
\end{document}
