\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}Higher Certificate, Paper I, 2005. Question 5

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{framed}
5. (i) A social scientist is conducting a survey of drug-taking among students. Her
survey is based on face-to-face interviews with a random sample of n students
from Wackford Squeers University, y of whom tell her that they take drugs.
Assume that the true proportion of drug-takers in the population being sampled
is p, that the responses are truthful, and that the number in the sample who take
drugs follows the binomial distribution $B(n, p)$. Write down the likelihood
function for these data and find the maximum likelihood estimator (MLE) of $p$,
$\hat{p}$ say. Also write down Var( pˆ ) and hence obtain an estimate of the standard
error of $\hat{p}$ , $SE(\hat{p})$ say. Calculate the values of $\hat{p}$ and $SE(\hat{p})$, given that
n = 100 and y = 20.
\end{framed}
%----------------------------------------------------------------%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{enumerate}[(a)]
\item We have Y ~ B(n, p), so ( ) ( ) ! (1 )

 probability mass function:

\[{\displaystyle f(y,n,p)=\Pr(k;n,p)=\Pr(Y=y)={\binom {n}{y}}p^{y}(1-p)^{n-y}} \]
for $y = 0, 1, 2, ..., n$, and $0 < p < 1$. 
\begin{itemize}
\item
The likelihood L is simply P(Y = y).
\item
Hence \[\log L = \mbox{constant} + y \log p + (n − y)\log(1− p).\]
\item 
\[ \therefore \frac{d \log L}{dp} = \frac{y}{p} - \frac{n-y}{1-p}\]
which on setting equal to zero gives that the maximum
likelihood estimate is pˆ y
n
= . 

\item Consideration of
${ \displaystyle \frac{d^2(\log L)}{dp^2} } $ confirms that this is a
maximum.

\item 
\begin{eqnarray*}
\operatorname{Var}(\hat{p}) &=& 
\frac{1}{n^2}\operatorname{Var}(Y)\\
&=& \frac{1}{n^2}np(1-p)\\
\frac{p(1-p)}{n}\\
\end{eqnarray*}

\item We may estimate p by $\hat{p}$ in this and
thus obtain an estimate of the standard error of $\hat{p}$ as 
\[SE (\hat{p}) = \sqrt{\frac{\hat{p} \times (1-\hat{p}) }{n}.\]
\item 
For $n = 100$ and $y = 20$, we have $\hat{p} = 0.2$ and 
\[SE( \hat{p} )  =  \sqrt{ \frac{0.2 \times 0.8}{100} } = 0.04\]
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{framed}
(ii) A statistician now advises her that, because of the sensitivity of the question of
drug-taking, some students may not answer truthfully, so causing bias.
\begin{itemize}
    \item He
suggests using an anonymising device to encourage truthful answers. 
\item He
provides a biased coin with faces labelled "takes drugs" and "does not take
drugs", which show with respective probabilities $0.75$ and $0.25$ when the coin
is tossed.
\item A second random sample, also of size $n$, is selected independently of the first.
 
\item Each of the n students is asked to toss the coin, unseen by the interviewer, and
to answer "yes" if he/she is in the group indicated by the coin and "no"
otherwise.
\end{itemize}
Assuming truthful responses, show that the probability of a "yes", $\theta$ say, is
given by
$\theta = 0.25 + 0.5p$ .
Assuming further that the number of students answering "yes", Z say, is
distributed $B(n, \theta )$ with observed value $z$, write down the MLE of $\theta$, $\hat{\theta}$ say,
and an estimate of its standard error. 


Use the relationship between $\theta$ and $p$ to
deduce the MLE of $p$, $\tilde{p}$ ~ say, and $SE(\tilde{p} )$. Calculate the values of $\tilde{p}$ and
$SE(\tilde{p})$, given that $n = 100$ and $z = 45$.
\end{framed}
%----------------------------------------------------------------%

\item P("yes") is given by P(takes drugs and coin shows "takes drugs") + P(does not
take drugs and coin shows "does not take drugs").
\begin{itemize}
    \item Hence $\theta = P("\mbox{yes}") = 0.75p + 0.25(1 – p) = 0.25 + 0.5p.$
    \item $Z \sim B(n, \theta )$, so from part (i) we have $$\hat{theta} = z/n$ and $SE(\hat{\theta}) = \hat{\theta} (1-\hat{\theta})/ n$ .
    \item We have $p = 2 \theta - 1/2$, so the MLE of p is $\tilde{p} = 2 \hat{\theta} - 1/2$
    \item Thus 
\begin{eqnarray*}
S.E.( \tilde{p} ) &=& 2 SE(\hat{\theta}) \\ 
&=& 2 \hat{\theta} (1-\hat{\theta})/ n .
\end{eqnarray*}
    \item For n = 100 and z = 45, we have $$\hat{theta} = 0.45$ and so $\tilde{p}= 0.4$ and $SE( \tilde{p} ) = 0.0995$ .
\end{itemize}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\begin{framed}
(iii) A journalist comparing the results of the two surveys says that the first survey
is more reliable. Do you agree? Why, or why not?

\end{framed}
%----------------------------------------------------------------%
\item Both p and the standard error are estimated to be larger by the second survey.
\begin{itemize}
    \item The larger p is plausible, as some people are likely not to admit to taking drugs when
asked directly as in the first survey.
    \item There is a much better expectation of truthful
answers in the second survey.
    \item We should not claim that the first is better just because
the SE is smaller; the first survey is likely to be biased. 
    \item Is it clear what the journalist
means by "reliable"?
\end{itemize}

\end{enumerate}

\end{document}
