\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}



\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

\noindent Of the adult population in a large city, 60\% favour a new leisure complex, 30\% oppose it and 10\% are indifferent.  A random sample of 4 adults is taken from the population and their opinions on the new complex are noted. 
 
\begin{framed} 
Find the probability that 

\begin{enumerate}[(i)]
\item all four think alike, 
\item  none of the four is opposed to the complex, 
\item all three opinions (in favour, oppose, indifferent) are represented in the sample, 
\item  all four are in favour of the complex, if it is given that none of the four is opposed. (10) 
\end{enumerate} 
\end{framed} 

 
 



%-Higher Certificate, Paper I, 2004. Question 1
\begin{enumerate}

\item (a)
\[P(\mbox{all four favour the complex}) = (0.6)^4.\]
\[P(\mbox{all four oppose the complex}) = (0.3)^4.\]
\[P(\mbox{all four are indifferent}) = (0.1)4.\]
So 

\begin{eqnarray*}
P(\mbox{all four think alike}) 
&=& (0.6)^4 + (0.3)^4 + (0.1)^4 \\ &=& 0.1378.
\end{eqnarray*}


\item  P(an individual is not opposed) = 0.6 + 0.1 = 0.7.
So 

\begin{eqnarray*}
P(\mbox{none of the four is opposed}) &=& (0.7)^4\\ &=& 0.2401
\end{eqnarray*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\item  Possible favourable results are FFOI, FOOI, FOII, in any order.
\begin{itemize}
\item $P(FFOI) = (0.6)^2(0.3)(0.1) = 0.0108$
\item $P(FOOI) = (0.6)(0.3)^2(0.1) = 0.0054$
\item $P(FOII) = (0.6)(0.3)(0.1)^2 = 0.0018$
\end{itemize}

Each result can be arranged in 4!
2!1!1!
= 12 ways.

So overall probability is 12(0.0108 + 0.0054 + 0.0018) = 0.216.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\item 

\begin{itemize}
\item From (a), P(all four in favour) = (0.6)^4 = 0.1296.
\item From (b), P(none
opposed) = 0.2401. 
\item So the required conditional probability is
0.1296/0.2401 = 0.5398.
\end{itemize} 


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{framed}
(ii) State the expectation and variance of the number in the sample who are in favour of the complex. (3) 
\end{framed}
\item The number in favour is binomially distributed with n = 4 and p = 0.6. 
\begin{itemize}
    \item So the
expectation (mean) is $4 \times 0.6 = 2.4$ 
\item the variance is $4 \times 0.6 \times 0.4 = 0.96$.
\end{itemize}

\begin{framed} In this city, one quarter of adults are classified as "young" (age < 30) and three-quarters are "older" (aged at least 30).  You are told that 12\% of young adults oppose the complex;  deduce the proportion of older adults who are opposed.  
\end{framed}

\item P(opposed) = P(opposedyoung)P(young) + P(opposedolder)P(older)
= (0.12 × 0.25) + (p × 0.75)
where p = P(opposedolder). But we are given that P(opposed) = 0.3. Hence
p = 0.36.


 
 
\begin{framed} Given that the sample consists of one young adult and three older adults, find the probability that exactly one member of the sample opposes the complex. 
\end{framed}
\item In samples of one "young" and three "olders",
\begin{eqnarray*}
P(exactly one opposes) &=& P("young" opposes, "olders" do not)
 + P("young" does not oppose, one "older" opposes)\\
&=& {(0.12)(0.64)3} + {3(0.88)(0.36)(0.64)2}\\ &=& 0.03146 + 0.38928\\
&=& 0.4207.
\end{eqnarray*}
\end{enumerate}
\end{document}
