\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}

\maketitle

\section{Introduction}
\begin{enumerate}
    \item (i) H0: cerebral blood flow and senile dementia are independent.
H1: they are not independent.
Calculate expected frequencies in the 2-way table:
OBS Blood Flow Normal Reduced EXP N R
Dementia No 92 28 120 No 86 34
Y ES 80 40 120 Y ES 86 34
172 68 240 172 68
Â2
(1) = 62( 2
86 + 2
34 ) = 72£ 120
86£34 = 2:955,n.s.,providing no statistical evidence for rejecting
H0.
\item  For a matched case-control study, McNemar’s test is required, which has greater power
of discrimination because of the matching.
CONTROL
N R
CASE N 74 6
R 18 22
Â2
(1) = (6¡18)2
6+18 = 144
24 = 6:00, so H0 is rejected.
\end{enumerate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Definition}
The test is applied to a 2 × 2 contingency table, which tabulates the outcomes of two tests on a sample of n subjects, as follows. 

Test 2 positive
Test 2 negative
Row total 
Test 1 positive
a
b
a + b 
Test 1 negative
c
d
c + d 
Column total
a + c
b + d
n 
The null hypothesis of marginal homogeneity states that the two marginal probabilities for each outcome are the same, i.e. pa + pb = pa + pc and pc + pd = pb + pd. 
Thus the null and alternative hypotheses are[1] 
\[{\displaystyle {\begin{aligned}H_{0}&:~p_{b}=p_{c}\\H_{1}&:~p_{b}\neq p_{c}\end{aligned}}} \]

Here pa, etc., denote the theoretical probability of occurrences in cells with the corresponding label. 
The McNemar test statistic is: 
χ 2 = ( b − c ) 2 b + c . \[{\displaystyle \chi ^{2}={(b-c)^{2} \over b+c}.} \]

Under the null hypothesis, with a sufficiently large number of discordants (cells b and c), 
${\displaystyle \chi ^{2}} $
 has a chi-squared distribution with 1 degree of freedom. If the 
${\displaystyle \chi ^{2}} $
 result is significant, this provides sufficient evidence to reject the null hypothesis, in favour of the alternative hypothesis that pb ≠ pc, which would mean that the marginal proportions are significantly different from each other. 
\end{document}