\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}
\begin{document}
\begin{enumerate}
\item A suitable N.H. is that the population distributions are the same with A.H.that
they are different. On the N.H. the number of instance where the new types will give a
higher figure is Binomial(10; p = y2).
there are 8 differences in favor of the original,and 2 for the new.
In B(10; 1
2 ), P(2 or less) = (1 + 10 + 45)(1
2 )10 = 0:0547 not significant. so the N.H. is
not rejected.

%%%%%%%%%%%%%%%%%%%%%%%
\item For a Wilcoxon test the N.H. is that population distributions are identical,and
the A.H is that they differ in location.
The actual difference(original-new) for each car are: +0.4,+0.3,-0.6,+0.8,+0.2,-0.1,+0.3+0.4,+1.1,
+0.2 and the ranks are 61
2 ; 41
2 ; 8; 9; 21
2 ; 1; 41
2 ; 61
2 ; 10; 21
2 : 

\begin{itemize}
    \item The sums of ranks are T¡ =
8 + 1 = 9; T+ = 46 the smaller of these is compared with the critical value for n=10,
which is T = 8; T¡ > 8 so the N.H. is not rejected.
\item However, the result in (ii) is nearer to significance than that in(i)).
\item If we could assume that the differences for the 10 cars followed a normal distribution,
then a paired t-test could be applied.  \itemIt would, if valid, be were powerful than either of
those above.  
\end{itemize}
\end{enumerate}
\end{document}
