\documentclass[a4paper,12pt]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}

%\usepackage{bigints}
\usepackage{vmargin}

% left top textwidth textheight headheight

% headsep footheight footskip

\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}

\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}

\begin{document}
Higher Certificate, Paper III, 2006. Question 7
(i)
Solution continued on next page
There is some curvature in all these plots; log y on x is slightly "straighter" than y on x. Using log x looks worst. So use log y on x.
(ii) (c) has the highest R2 – though all are good.
(c) also has the highest t values for the coefficients – though again all are good.
(c) also has the lowest residual variance relative to the mean response.
(c) is the only one without a "large" residual.
(c) appeared (marginally) the best plot.
It does not seem sensible to regress log y on log x; it looks as if this would increase the curvature.
(iii) Using (c), we have log y = 5.41 – 0.322x, so
()5.410.3220.322exp5.410.322223.63xxyxee−−=−== .
(iv) From the expression in part (iii), inserting x = 5, 1.61223.63ye−= = 44.7. This is in hundreds per square kilometre. So the estimate is 4470 per square kilometre.
(v) Prediction within the range of the data may be adequate, except perhaps near the upper end because of the tendency for curvature there. Extrapolation to values of x outside the data will, for similar reasons, be unreliable, and the linear model is likely to underestimate density. Where is the next city or town centre? Interaction with that is very likely unless it is a long distance away. There may also be directional effects, i.e. densities changing more or less slowly according to the direction from the centre.
 \end{enumerate}
 \end{document}
 
