\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}
\begin{document}

\begin{enumerate}
\item For type A, min = 171; lower quartile, q = 396:5; median, M = 1
2(568+795) =
681:5; upper quartile, Q = 1158; max = 2415.
For B, min = 212; q = 298:5; M = 447:5; Q = 823:5; max = 1678.

\begin{itemize}
\item The two distributions are distinctly skew, since the medians are not in the
middle of the boxes made by the quartiles, and also the upper whiskers are
very long. 
\item The variability in the distributions appears not to be the same
either.
\end{itemize}

\item The t-test requires symmetry (strictly normality) of sets of data and, at least
approximately, the same variance. Since neither of these seems very likely
in the populations from which the samples were drawn, a Mann - Whitney
test is preferred. This requires data to be of similar shape, but that is more
24
reasonable. The Null Hypothesis will be that the populations have the same
median values. The ranks of Type A are: \[\{1, 4, 8, 9, 13, 14, 16, 17, 20, 21,
25, 28, 30, 32, 33, 34, 35, 37, 38, 40\}\]; and of type B: \[\{2, 3, 5, 6, 7, 10, 11, 12,
15, 18, 19, 22, 23, 24, 26, 27, 29, 31, 36, 39.\}\]

\begin{itemize}
\item Rank sums are: A, 455; B, 365. [check: sum = 820 = 1
2 ¢ 40 ¢ 41]. 
\item The mean
of all ranks is 410, and the normal approximation to rank sum has variance
1
12 ¢ 20 ¢ 20 ¢ 41, i.e. s.d.= 36:97.
(This form of the test is usually called Wilcoxan’s Rank Sum test.)
\item Hence r = 455¡410
36:97 = 1:22 is approximately N(0; 1); the value is not significant,
so there is no evidence that medians differ.
\end{itemize}
\end{enumerate}
\end{document}
