\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}
\begin{document}
Higher Certificate, Paper II, 2003. Question 4
\begin{enumerate}[(a)]
\item (i) A sign test is appropriate, the null hypothesis being that A and V are equally
likely to be the greater. Hence the number of As is binomial with parameters
10 and ½, as is the number of Vs, if the null hypothesis is true. The alternative
hypothesis claims that V is greater. A one-sided test is therefore needed.
The observed results are nA = 3, nB = 7.
If the null hypothesis is true, we have
( ) 10 10
10 10 10 7 1 1 120 45 10 1 0.172
7 8 9 2 2 V P n
       + + + ≥ =   +   +   +  = =
      
.
There is not enough evidence to reject the null hypothesis.
The sign test is not very powerful. The sample size here (8) is not really large
enough for its effective use.
\item  (a) The data provide information which the sign test would not use,
namely that measuring the change on a numerical scale. As well as the
sign of the change we should use the size. A Wilcoxon signed-rank
test is suitable. Difference A – B are as follows, and the absolute
values of the differences are ranked in size order, with tied values
given their average ranking.
(1) (2) (3) (4) (5) (6) (7) (8) (9) (10) (11) (12)
7 –2 6 4 24 15 –5 1 3 5 –7 –1
Rank 9½ 3 8 5 12 11 6½ 1½ 4 6½ 9½ 1½
The null hypothesis is that aural memory scores are not altered by
coaching, the alternative hypothesis is that coaching leads to
improvement. A one-sided test is required.
The sum of ranks of negative values is T– = 20½ and of positive values
is T+ = 57½. We use T = min(T–, T+) = 20½. Tables for n = 12 give 17
as the critical value for a one-sided 5% test, so there is not enough
evidence to reject the null hypothesis.
(b) A paired-samples t test using the differences would be appropriate if
the distribution of differences appeared to be approximately Normal.
The two large values for (5) and (6) make this unlikely, both being on
the same side (+). It would be unwise to use the t test in this case.
\end{enumerate}
\end{document}
