\documentclass[a4paper,12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{eurosym}
\usepackage{vmargin}
\usepackage{amsmath}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{enumerate}
\usepackage{multicol}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{listings}
\usepackage{framed}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{chngpage}
%\usepackage{bigints}

\usepackage{vmargin}
% left top textwidth textheight headheight
% headsep footheight footskip
\setmargins{2.0cm}{2.5cm}{16 cm}{22cm}{0.5cm}{0cm}{1cm}{1cm}
\renewcommand{\baselinestretch}{1.3}

\setcounter{MaxMatrixCols}{10}
\begin{document}

Higher Certificate, Paper II, 2003. Question 8

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}[ht!]
 
\centering
 
\begin{tabular}{|p{15cm}|}
 
\hline  
8. (i) Using examples to illustrate your answers, discuss the uses of the F distribution in statistical methods. (8) 

\\ \hline
  
\end{tabular}

\end{table}





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\begin{enumerate}[(a)]
\item (i) The F distribution with m and n degrees of freedom is the ratio of two
independent chi-squared distributions divided by their numbers of degrees of
freedom:
2
, 2
χ
χ
m
m n
n
F m
n
= .
Thus if estimates of variance s1
2 and s2
2 are obtained from each of two
independent samples of Normally distributed data, (m+1) and (n+1) items in
the samples respectively, the null hypothesis that the underlying population
variances σ 1
2 and σ 2
2 are equal can be tested by referring s1
2/s2
2 to Fm,n. A
confidence interval for σ 1
2/σ 2
2 can also be found.
\begin{itemize}
    \item In analysis of variance, sums of squares will have χ2 distributions if Normality
is assumed. A suitable null hypothesis is set up, such as that a regression
coefficient is zero. 
\item The appropriate sum of squares is then compared with the
residual sum of squares, using the "mean squares" so as to deal with the
numbers of degrees of freedom, using an F distribution.
\item If the null hypothesis
is true, both the mean squares will merely estimate experimental error so their
ratio has expected value 1, but if it is false the mean square corresponding to
the regression coefficient is expected to be larger. One-way and two-way (and
higher-way) analysis of variance for experimental designs uses the same
background theory.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\newpage
\begin{table}[ht!]
 
\centering
 
\begin{tabular}{|p{15cm}|}
 
\hline  

 (ii) A manufacturer of soft drinks needs to purchase a new bottle-filling machine.  The purchasing department has identified two machines, A and B, that are identical with respect to size, cost and convenience.  The deciding factor will be the variability in the amount of liquid delivered by each machine, with the machine having the lower variability being preferred.  Prior to purchase, the soft drinks manufacturer negotiates the chance to have each machine on one week's trial.  During this week, the normal production run is shared between the two machines, each being set to produce on average one litre of soft drink per bottle.  A random sample of bottles obtained from the production run of each machine is examined and the actual quantities of soft drink produced recorded in millilitres as follows. 
 
 
Machine A 
 1002   1002   1001   995   1000   1007   1001   1004   998   993   1002   999 998   999   1000   1001 
 
 
Machine B 
 1002   998   1001   1001   1000   999   1002   1000   1002   1000   1000   997 1001   1000   996   1001   1000   999   999   1003 
 
 
Using a suitable statistical test, compare the variabilities in the amounts delivered by the two machines and hence write a short report to the manufacturer stating your recommendations concerning the choice of purchase. (12) 

\\ \hline
  
\end{tabular}

\end{table}
\item  The null hypothesis is "σA
2 = σB
2", the same variability for both machines, and
the alternative hypothesis is "σA
2 ≠ σB
2".
Summary statistics for the two machines are:
2
1 1 1 A: n =16, x =1000.125, s =10.9167 (s = 3.304) ;
2
2 2 2 B : n = 20, x =1000.050, s = 2.9974 (s =1.731) .
s1
2/s2
2 = 3.64, refer to F15,19. This is significant at the 5\% level so the null
hypothesis can be rejected and we can decide to use B because the evidence is
that it is less variable than A.
A report should mention that both machines dispense, on average, just over
1000 ml. However, the data from the trial period show that A was
significantly more variable than B (in fact, inspection of the data shows two
rather outlying values in A, at 993 and 1007), so B is the one to buy.
\end{enumerate}
\end{document}
